---
title: "manu4_diffExp"
author: "Katja K"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}

library(DESeq2)
library(tidyverse)

```

# Metadata 

```{r metadata}

sampledata <- read.table("../seq_data/sample.data", row.names = 1, header=T)
sampledata$Time <- as.factor(sampledata$Time)

```

# Differential expression 
## model all
+ based on raw counts (filtered raw counts)
+ 2 countries (SE, ES), 2 treatments (C, D)
+ response of both countries to treatment
+ differences in the response between countries (interaction)

```{r diffExp overall, message=FALSE, warning=FALSE}

#unfiltered counts, ID level
#gene_counts_ID <- read.table("../../seq_data/enzID_counts_sort.tsv", header = T, row.names = 1, check.names = F)
gene_counts_ID <- read.table("../seq_data/enzID_filt_counts.tsv", header = T, row.names = 1, check.names = F)


set.seed(6282)
de_all <- DESeqDataSetFromMatrix(countData=ceiling(gene_counts_ID),colData=sampledata,design=~Country+Treatment)

de_all$Country <- as.factor(de_all$Country)
de_all$Treatment <- as.factor(de_all$Treatment)

design(de_all) <- ~ Country + Treatment + Country:Treatment
set.seed(2741)
de_all <- DESeq(de_all) 
resultsNames(de_all)

saveRDS(de_all, file = "../seq_data/de_all_model.rds")

```

## drought in Spain

```{r drought in ES, message=FALSE, warning=FALSE}

#The effect of drought in Spain (the main effect).
res = results(de_all, contrast=c("Treatment","D","C"))
res$padj[is.na(res$padj)] <- 1
summary(res)

# all genes
nrow(as.data.frame(res))
# only genes with padj <0.05
nrow(dplyr::filter(as.data.frame(res),padj<0.05))
#down regulated
nrow(dplyr::filter(as.data.frame(res),padj<0.05,(log2FoldChange)<0))
#up regulated
nrow(dplyr::filter(as.data.frame(res),padj<0.05,(log2FoldChange)>0))

res_signif <- dplyr::filter(as.data.frame(res),padj<0.05)
dim(res_signif)

# sort
res_signif <- res_signif[order(abs(res_signif$log2FoldChange),decreasing = T),] 
write.table(res_signif,"./results/DE_all_D_ES.tsv",row.names = TRUE, col.names = TRUE, sep = "\t")

```

## drought in Sweden

```{r drought in SE, message=FALSE, warning=FALSE}

#The effect of drought in Sweden
res = results(de_all, list(c("Treatment_D_vs_C","CountrySE.TreatmentD")))
res$padj[is.na(res$padj)] <- 1
summary(res)

# all genes
nrow(as.data.frame(res))
# only genes with padj <0.05
nrow(dplyr::filter(as.data.frame(res),padj<0.05))
#down regulated
nrow(dplyr::filter(as.data.frame(res),padj<0.05,(log2FoldChange)<0))
#up regulated
nrow(dplyr::filter(as.data.frame(res),padj<0.05,(log2FoldChange)>0))

res_signif <- dplyr::filter(as.data.frame(res),padj<0.05)
dim(res_signif)

# sort
res_signif <- res_signif[order(abs(res_signif$log2FoldChange),decreasing = T),] 
write.table(res_signif,"./results/DE_all_D_SE.tsv",row.names = TRUE, col.names = TRUE, sep = "\t")

```

## difference between ES and SE in C

```{r all ESvsSE control, message=FALSE, warning=FALSE}

res = results(de_all, contrast=c("Country","ES","SE"))
res$padj[is.na(res$padj)] <- 1
summary(res)

# all genes
nrow(as.data.frame(res))
# only genes with padj <0.05
nrow(dplyr::filter(as.data.frame(res),padj<0.05))
nrow(dplyr::filter(as.data.frame(res),padj<0.05,(log2FoldChange)<0))
nrow(dplyr::filter(as.data.frame(res),padj<0.05,(log2FoldChange)>0))

res_signif <- dplyr::filter(as.data.frame(res),padj<0.05)

# sort
res_signif <- res_signif[order(abs(res_signif$log2FoldChange),decreasing = T),] 
write.table(res_signif,"./results/DE_all_ESvsSE_C.tsv",row.names = TRUE, col.names = TRUE, sep = "\t")

```

## difference between ES and SE in D

```{r all ESvsSE drought, message=FALSE, warning=FALSE}

res = results(de_all, list(c("Country_SE_vs_ES", "CountrySE.TreatmentD")))
res$padj[is.na(res$padj)] <- 1
summary(res)

# all genes
nrow(as.data.frame(res))
# only genes with padj <0.05
nrow(dplyr::filter(as.data.frame(res),padj<0.05))
nrow(dplyr::filter(as.data.frame(res),padj<0.05,(log2FoldChange)<0))
nrow(dplyr::filter(as.data.frame(res),padj<0.05,(log2FoldChange)>0))

res_signif <- dplyr::filter(as.data.frame(res),padj<0.05)

# sort
res_signif <- res_signif[order(abs(res_signif$log2FoldChange),decreasing = T),] 
write.table(res_signif,"./results/DE_all_ESvsSE_D.tsv",row.names = TRUE, col.names = TRUE, sep = "\t")

```

## response across countries (interaction)

```{r all interaction, message=FALSE, warning=FALSE}

res = results(de_all, name="CountrySE.TreatmentD")
res$padj[is.na(res$padj)] <- 1
summary(res)

# all genes
nrow(as.data.frame(res))
# only genes with padj <0.05
nrow(dplyr::filter(as.data.frame(res),padj<0.05))
nrow(dplyr::filter(as.data.frame(res),padj<0.05,(log2FoldChange)<0))
nrow(dplyr::filter(as.data.frame(res),padj<0.05,(log2FoldChange)>0))

res_signif <- dplyr::filter(as.data.frame(res),padj<0.05)

# sort
res_signif <- res_signif[order(abs(res_signif$log2FoldChange),decreasing = T),] 
write.table(res_signif,"./results/DE_all_interaction.tsv",row.names = TRUE, col.names = TRUE, sep = "\t")

```

## separate DE
+ based on raw counts
+ for each country separately
+ treatment (2 levels)

### Spain

```{r diffExp Spain, message=FALSE, warning=FALSE}

es_meta <- sampledata %>%
  filter(Country == "ES")
es_meta$Farm <- factor(es_meta$Farm)

counts_ES <- gene_counts_ID[-c(11:20,31:40)]

set.seed(6282)
de_spain <- DESeqDataSetFromMatrix(countData=ceiling(counts_ES),colData=es_meta,design=~Treatment)

design(de_spain) <- ~ Treatment 
set.seed(2741)
de_spain <- DESeq(de_spain) 
resultsNames(de_spain)

saveRDS(de_spain, file = "../seq_data/de_ES_model.rds")

```

```{r}

d_spain <- readRDS("../seq_data/de_ES_model.rds")

res = results(d_spain, contrast=c("Treatment","D","C"))
res$padj[is.na(res$padj)] <- 1
summary(res)

# all genes
nrow(as.data.frame(res))
# only genes with padj <0.05
nrow(dplyr::filter(as.data.frame(res),padj<0.05))
nrow(dplyr::filter(as.data.frame(res),padj<0.05,(log2FoldChange)<0))
nrow(dplyr::filter(as.data.frame(res),padj<0.05,(log2FoldChange)>0))

res_signif <- dplyr::filter(as.data.frame(res),padj<0.05)
nrow(res_signif)

# sort
res_signif <- res_signif[order(abs(res_signif$log2FoldChange),decreasing = T),] 
write.csv(res_signif,"./results/DE_ES_drought.csv",row.names=T)

```

### Sweden

```{r diffExp Sweden, message=FALSE, warning=FALSE}

se_meta <- sampledata %>%
  filter(Country == "SE")
se_meta$Farm <- factor(se_meta$Farm)

counts_SE <- gene_counts_ID[-c(1:10,21:30)]

set.seed(6282)
de_sweden <- DESeqDataSetFromMatrix(countData=ceiling(counts_SE),colData=se_meta,design=~Treatment)

design(de_sweden) <- ~ Treatment
set.seed(2741)
de_sweden <- DESeq(de_sweden) 
resultsNames(de_sweden)

saveRDS(de_sweden, file = "../seq_data/de_SE_model.rds")

```

```{r}

d_sweden <- readRDS("../seq_data/de_SE_model.rds")

res = results(d_sweden, contrast=c("Treatment","D","C"))
res$padj[is.na(res$padj)] <- 1
summary(res)

# all genes
nrow(as.data.frame(res))
# only genes with padj <0.05
nrow(dplyr::filter(as.data.frame(res),padj<0.05))
nrow(dplyr::filter(as.data.frame(res),padj<0.05,(log2FoldChange)<0))
nrow(dplyr::filter(as.data.frame(res),padj<0.05,(log2FoldChange)>0))

res_signif <- dplyr::filter(as.data.frame(res),padj<0.05)
nrow(res_signif)

# sort
res_signif <- res_signif[order(abs(res_signif$log2FoldChange),decreasing = T),] 
write.csv(res_signif,"./results/DE_SE_drought.csv",row.names=T)

```


```{r}

sessionInfo()
```
