---
title: "manuscript_4_sequencing_overview"
author: "Katja K"
output: 
  rmarkdown::html_document:
    toc: true
    depth: 3
    number_sections: true  
    theme: united  
    highlight: tango
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r wrap-hook, include=FALSE}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```

```{r packages, include=FALSE}

library(ggplot2)
library(vegan)
library(DESeq2)
library(edgeR)
library(cowplot)
library(ggrepel)
library(dplyr)
library(wesanderson)

```

# Metadata 

```{r metadata, results='hide'}

sampledata <- read.table("../../seq_data/sample.data", row.names = 1, header=T)
sampledata$Time <- as.factor(sampledata$Time)
sampledata$Farm <- as.factor(sampledata$Farm)

```

# Counts EZfam
## filterByExpr

```{r filter counts EZfam, warning=FALSE, message=FALSE}

gene_counts_EZfam <- read.table("../../seq_data/fam_counts.tsv", header = T, row.names = 1, check.names = F)
dim(gene_counts_EZfam)
sum(gene_counts_EZfam)

EZfam_dgelist <- DGEList(gene_counts_EZfam)
EZfam_dgelist <- calcNormFactors(EZfam_dgelist,method=c("TMM")) #trimmed mean of M-values

table(rowSums(EZfam_dgelist$counts==0)==40)

#automatic way to filter genes
keep.exprs <- filterByExpr(EZfam_dgelist)

#number of genes left after filtering
EZfam_dgelist <- EZfam_dgelist[keep.exprs,, keep.lib.sizes=FALSE]
dim(EZfam_dgelist)

sum(EZfam_dgelist$counts) 
307301666/307311474  

``` 

## order

```{r order EZfam, message=FALSE, warning=FALSE, echo=FALSE}

all(colnames(EZfam_dgelist) == rownames(sampledata))
sort(colnames(EZfam_dgelist))
sort(rownames(sampledata))

EZfam_dgelist_ord <- match(rownames(sampledata), colnames(EZfam_dgelist))
EZfam_ord <- EZfam_dgelist[,EZfam_dgelist_ord]
all(rownames(sampledata) == colnames(EZfam_ord))

EZfam_dgelist <- EZfam_ord 

write.table(EZfam_dgelist$counts, "../../seq_data/EZfam_filt_counts.tsv",row.names = TRUE, col.names = TRUE, sep = "\t" )

```

## VST EZfam

```{r VST EZfam, message=FALSE, warning=FALSE}

EZfam_dgelist <- read.table("../../seq_data/EZfam_filt_counts.tsv", header = T, row.names = 1, check.names = F)

set.seed(6253)
dd_fam <- DESeqDataSetFromMatrix(countData=EZfam_dgelist,colData=sampledata,design=~Country+Treatment+Time)

vst_EZfam <- varianceStabilizingTransformation(dd_fam)

p1_EZfam_country <- plotPCA(vst_EZfam, "Country")
p2_EZfam_treat <- plotPCA(vst_EZfam, "Treatment")
p3_EZfam_time <- plotPCA(vst_EZfam, "Time")
p4_EZfam_farm <- plotPCA(vst_EZfam, "Farm")
plot_grid(p1_EZfam_country, p2_EZfam_treat, p3_EZfam_time, p4_EZfam_farm)

```

## Permanova

```{r permanova EZfam, message=FALSE, warning=FALSE}

vst_EZfam_counts <- assay(vst_EZfam)
vst_EZfam_counts <- t(vst_EZfam_counts)

set.seed(5392)
perm1_EZfam <- adonis(vst_EZfam_counts ~ Country+Treatment+Time+Country:Treatment+Country:Time+Treatment:Time+Country:Treatment:Time, 
                   data=sampledata, permutations=999, strata = sampledata$Farm)
perm1_EZfam

```
***Strong country effect, interaction between country:time***

## PCA VST EZfam

```{r PCA EZfam VST, message=FALSE, warning=FALSE}

#samples should be rows,variables columns
vst_EZfam_pca <- prcomp(vst_EZfam_counts)

PCA_vst_EZfam <- data.frame(vst_EZfam_pca$x, Country=sampledata$Country, Farm=sampledata$Farm, Time=sampledata$Time, Treat=sampledata$Treatment, Cou_Treat=sampledata$Cou_Treat)

PCA_vst_EZfam$Cou_Treat <- factor(PCA_vst_EZfam$Cou_Treat, levels = c("ES_C", "ES_D", "SE_C", "SE_D"), labels = c("Spain + C", "Spain + D", "Sweden + C", "Sweden + D"))

ggplot() + geom_point(data=PCA_vst_EZfam, mapping=aes(x=PC1, y=PC2, color = Cou_Treat, shape=Time), size = 10,alpha=0.8) +
    geom_vline(xintercept = 0, linetype=2) +
    geom_hline(yintercept = 0, linetype=2) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank()) + labs(colour = "Country") +  
    scale_color_manual(values=c("#4e032c","#b10764","#034e4b","#069891")) + labs(colour = "Country + drought treatment", shape= "Sampling time") + guides(color = guide_legend(order = 1),
         shape = guide_legend(order = 2)) +
  theme(axis.text = element_text(size = 12)) + 
    theme(axis.title = element_text(size = 12)) +
    theme(legend.title = element_text(size = 12)) + 
    theme(legend.text = element_text(size = 12))

```

## Batch correction EZfam
+ correction on countries
+ PERMANOVA for batch

```{r batch EZfam, message=FALSE,warning=FALSE}

vst_fam_remCOUNTRY <- vst_EZfam
assay(vst_fam_remCOUNTRY) <- limma::removeBatchEffect(assay(vst_fam_remCOUNTRY), vst_fam_remCOUNTRY$Country)

p1_EZfam_country_corr <- plotPCA(vst_fam_remCOUNTRY, "Country")
p2_EZfam_treat_corr <- plotPCA(vst_fam_remCOUNTRY, "Treatment")
p3_EZfam_time_corr <- plotPCA(vst_fam_remCOUNTRY, "Time")
p4_EZfam_farm_corr <- plotPCA(vst_fam_remCOUNTRY, "Farm")

plot_grid(p1_EZfam_country_corr, p2_EZfam_treat_corr, p3_EZfam_time_corr, p4_EZfam_farm_corr)

#permanova
vst_fam_remCOUNTRY_df <- as.data.frame(assay(vst_fam_remCOUNTRY))
vst_fam_remCOUNTRY_df <- t(vst_fam_remCOUNTRY_df)

set.seed(4364)
adonis(vst_fam_remCOUNTRY_df ~ Treatment, 
                     data=sampledata, permutations=999, strata = sampledata$Farm) 

```
***No drought effect***

# Counts EZ IDs
## filterByExpr

```{r filter counts EZ IDs, warning=FALSE, message=FALSE}

gene_counts_ID <- read.table("../../seq_data/enzID_counts.tsv", header = T, row.names = 1, check.names = F)
dim(gene_counts_ID)
sum(gene_counts_ID)

EZ_ID_dgelist <- DGEList(gene_counts_ID)
EZ_ID_dgelist <- calcNormFactors(EZ_ID_dgelist,method=c("TMM")) #trimmed mean of M-values

table(rowSums(EZ_ID_dgelist$counts==0)==40)

#automatic way to filter genes
keep.exprs <- filterByExpr(EZ_ID_dgelist)

#number of genes left after filtering
EZ_ID_dgelist <- EZ_ID_dgelist[keep.exprs,, keep.lib.sizes=FALSE]
dim(EZ_ID_dgelist)

sum(EZ_ID_dgelist$counts) 
282917374/307311474  

``` 

## order

```{r order ID level, message=FALSE, warning=FALSE, echo=FALSE}

all(colnames(EZ_ID_dgelist) == rownames(sampledata))
sort(colnames(EZ_ID_dgelist))
sort(rownames(sampledata))

geneID_dgelist_ord <- match(rownames(sampledata), colnames(EZ_ID_dgelist))
geneID_ord <- EZ_ID_dgelist[,geneID_dgelist_ord]
all(rownames(sampledata) == colnames(geneID_ord))

EZ_ID_dgelist <- geneID_ord

write.table(EZ_ID_dgelist$counts, "../../seq_data/enzID_filt_counts.tsv",row.names = TRUE, col.names = TRUE, sep = "\t" )

```

## VST EZ ID

```{r VST EZ ID, message=FALSE, warning=FALSE}

EZ_ID_dgelist <- read.table("../../seq_data/enzID_filt_counts.tsv", header = T, row.names = 1, check.names = F)

set.seed(6359)
dd_ID <- DESeqDataSetFromMatrix(countData=EZ_ID_dgelist,colData=sampledata,design=~Country+Treatment+Time)

vst_ID <- varianceStabilizingTransformation(dd_ID)

p1_ID_country <- plotPCA(vst_ID, "Country")
p2_ID_treat <- plotPCA(vst_ID, "Treatment")
p3_ID_time <- plotPCA(vst_ID, "Time")
p4_ID_farm <- plotPCA(vst_ID, "Farm")
plot_grid(p1_ID_country, p2_ID_treat, p3_ID_time, p4_ID_farm)

```

## Permanova

```{r permanova ID, message=FALSE, warning=FALSE}

vst_ID_counts <- assay(vst_ID)
vst_ID_counts <- t(vst_ID_counts)

set.seed(5392)
perm1_ID <- adonis(vst_ID_counts ~Country+Treatment+Time+Country:Treatment+Country:Time+Treatment:Time+Country:Treatment:Time, 
                   data=sampledata, permutations=999, strata = sampledata$Farm)
perm1_ID

```
***Country and drought effect***

## PCA VST IDs

```{r PCA VST ID, warning=FALSE, message=FALSE}

ntop <- 500
rv <- rowVars(assay(vst_ID))
select <- order(rv, decreasing = TRUE)[seq_len(min(ntop, length(rv)))]
mat <- t( assay(vst_ID)[select, ] )

pca<-prcomp(mat)
#pca <- as.data.frame(pca$x)
#pca$PC1

eigs <- pca$sdev^2
eigs[1] / sum(eigs)
eigs[2] / sum(eigs)

PCA_vst_ID <- data.frame(pca$x, Country=sampledata$Country, Farm=sampledata$Farm, Time=sampledata$Time, Treat=sampledata$Treatment, Cou_Treat=sampledata$Cou_Treat)

PCA_vst_ID$Cou_Treat <- factor(PCA_vst_ID$Cou_Treat, levels = c("ES_C", "ES_D", "SE_C", "SE_D"), labels = c("Spain + C", "Spain + D", "Sweden + C", "Sweden + D"))
PCA_vst_ID$Country <- factor(PCA_vst_ID$Country, levels = c("ES", "SE"), labels = c("Spain", "Sweden"))

ggplot() + geom_point(data=PCA_vst_ID, mapping=aes(x=PC1, y=PC2, color = Country, shape=Treat), size = 10,alpha=0.8) +
    geom_vline(xintercept = 0, linetype=2) +
    geom_hline(yintercept = 0, linetype=2) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank()) + xlab("PC1 (31% variance)") +  ylab("PC2 (14% variance)") +
    scale_color_manual(values=c("#79402E","#46acc8")) + labs(colour = "Country", shape= "Drought treatment") + guides(color = guide_legend(order = 1),
         shape = guide_legend(order = 2)) +
  theme(axis.text = element_text(size = 12)) + 
    theme(axis.title = element_text(size = 12)) +
    theme(legend.title = element_text(size = 12)) + 
    theme(legend.text = element_text(size = 12))

ggsave("../figures/PCA_overall_enzID_batch.jpeg", width = 20, height = 15, units = "cm")

```

# Only T4

```{r}

EZ_ID_dgelist_T4 <- EZ_ID_dgelist[,21:40]
sampledata_T4 <- sampledata %>%
  filter(Time == "T4")

set.seed(6359)
dd_ID_T4 <- DESeqDataSetFromMatrix(countData=EZ_ID_dgelist_T4,colData=sampledata_T4,design=~Country+Treatment)

vst_ID_T4 <- varianceStabilizingTransformation(dd_ID_T4)

p1_ID_country_T4 <- plotPCA(vst_ID_T4, "Country")
p2_ID_treat_T4 <- plotPCA(vst_ID_T4, "Treatment")
p4_ID_farm_T4 <- plotPCA(vst_ID_T4, "Farm")
plot_grid(p1_ID_country_T4, p2_ID_treat_T4, p4_ID_farm_T4)

```

# Separate countries
## Sweden

```{r SE, message=FALSE, warning=FALSE}

se_meta <- sampledata %>%
  filter(Country == "SE")
se_meta$Farm <- as.factor(se_meta$Farm)

se_ID <- EZ_ID_dgelist[,colnames(EZ_ID_dgelist) %in% rownames(se_meta)]

set.seed(6359)
se_dd <- DESeqDataSetFromMatrix(countData = se_ID,
                              colData = se_meta[match(colnames(se_ID), rownames(se_meta)), ],
                              design= ~  Treatment + Farm + Time)

se_vst <- varianceStabilizingTransformation(se_dd)

p1_SE_farm <- plotPCA(se_vst, "Farm")
p2_SE_treat <- plotPCA(se_vst, "Treatment")
p3_SE_time <- plotPCA(se_vst, "Time")
plot_grid(p1_SE_farm, p2_SE_treat, p3_SE_time)

#PERMANOVA
se_vst_counts <- assay(se_vst)
se_vst_counts <- t(se_vst_counts)
set.seed(4364)
adonis(se_vst_counts~ Treatment + Time, 
       data=se_meta,  strata=se_meta$Farm, permutations=999)

```
***Differences between farm sites***
***Effect of time***

### plot PCA

```{r plot PCA SE}

ntop <- 500
rv <- rowVars(assay(se_vst))
select <- order(rv, decreasing = TRUE)[seq_len(min(ntop, length(rv)))]
mat <- t( assay(se_vst)[select, ] )

pca_sweden<-prcomp(mat)

eigs <- pca_sweden$sdev^2
eigs[1] / sum(eigs)
eigs[2] / sum(eigs)

PCA_SE_ID <- data.frame(pca_sweden$x, Farm=se_meta$Farm, Time=se_meta$Time, Treat=se_meta$Treatment, Cou_Treat=se_meta$Cou_Treat)

#drought-time
ggplot() + geom_point(data=PCA_SE_ID, mapping=aes(x=PC1, y=PC2, color = Treat, shape=Time), size = 10,alpha=0.8) +
geom_vline(xintercept = 0, linetype=2) +
geom_hline(yintercept = 0, linetype=2) +
theme_bw() +
theme(panel.grid.major = element_blank(),
panel.grid.minor = element_blank()) +
scale_color_manual(values=c("#8f8f8f","#000000")) + labs(colour = "Drought", shape= "Sampling time") + guides(color = guide_legend(order = 1),
shape = guide_legend(order = 2)) +
theme(axis.text = element_text(size = 12)) +
theme(axis.title = element_text(size = 12)) +
theme(legend.title = element_text(size = 12)) +
theme(legend.text = element_text(size = 12))

#farm-drought
ggplot(PCA_SE_ID, aes(x=PC1, y=PC2, color = Farm, shape=Treat)) + geom_text_repel(aes(label=Time),max.overlaps = Inf, point.size = 8, segment.color=NA, show.legend = F,size=6) +
  geom_point(aes(fill = Farm, shape = Treat),size = 10,alpha=0.8) + scale_shape_manual(values = c(24,25)) +
  scale_fill_manual(values=wes_palette(n=5, name="Cavalcanti1")) +
  scale_color_manual(values=wes_palette(n=5, name="Cavalcanti1")) +
    geom_vline(xintercept = 0, linetype=2) +
    geom_hline(yintercept = 0, linetype=2) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank()) +
    labs(fill = "Farm", shape= "Drought", color="Farm") + 
  xlab("PC1 (36% variance)") + ylab("PC2 (19% variance)") +
  theme(axis.text = element_text(size = 12)) + 
    theme(axis.title = element_text(size = 12)) +
    theme(legend.title = element_text(size = 12)) + 
    theme(legend.text = element_text(size = 12))
  
#ggsave("../figures/PCA_SE_enzID.jpeg", width = 20, height = 15, units = "cm")

```

### batch SE
+ correction on farms

```{r SE batch correction, message=F, warning=F}

se_vst_remFP <- se_vst
assay(se_vst_remFP) <- limma::removeBatchEffect(assay(se_vst_remFP), se_vst_remFP$Farm)

p1_SE_FP_corr <- plotPCA(se_vst_remFP, "Farm")
p2_SE_treat_corr <- plotPCA(se_vst_remFP, "Treatment")
plot_grid(p1_SE_FP_corr, p2_SE_treat_corr)

#PERMANOVA 
se_vst_remFP_df <- as.data.frame(assay(se_vst_remFP))
se_vst_remFP_df <- t(se_vst_remFP_df)

set.seed(4364)
adonis(se_vst_remFP_df ~ Treatment, 
                     data=se_meta, permutations=999)

```
***No drought effect***

### plot batch SE

```{r SE batch plot}

ntop <- 500
rv <- rowVars(assay(se_vst_remFP))
select <- order(rv, decreasing = TRUE)[seq_len(min(ntop, length(rv)))]
mat <- t( assay(se_vst_remFP)[select, ] )

pca_sweden_batch<-prcomp(mat)

eigs <- pca_sweden_batch$sdev^2
eigs[1] / sum(eigs)
eigs[2] / sum(eigs)

PCA_SE_ID_batch <- data.frame(pca_sweden_batch$x, Farm=se_meta$Farm, Time=se_meta$Time, Treat=se_meta$Treatment, Cou_Treat=se_meta$Cou_Treat)

#drought-time
ggplot() + geom_point(data=PCA_SE_ID_batch, mapping=aes(x=PC1, y=PC2, color = Treat, shape=Time), size = 10,alpha=0.8) +
geom_vline(xintercept = 0, linetype=2) +
geom_hline(yintercept = 0, linetype=2) +
theme_bw() +
theme(panel.grid.major = element_blank(),
panel.grid.minor = element_blank()) +
scale_color_manual(values=c("#8f8f8f","#000000")) + labs(colour = "Drought", shape= "Sampling time") + guides(color = guide_legend(order = 1),
shape = guide_legend(order = 2)) +
theme(axis.text = element_text(size = 12)) +
theme(axis.title = element_text(size = 12)) +
theme(legend.title = element_text(size = 12)) +
theme(legend.text = element_text(size = 12))

#farm-drought
ggplot(PCA_SE_ID_batch, aes(x=PC1, y=PC2, color = Farm, shape=Treat)) + geom_text_repel(aes(label=Time),max.overlaps = Inf, point.size = 8, segment.color=NA, show.legend = F,size=6) +
  geom_point(aes(fill = Farm, shape = Treat),size = 10,alpha=0.8) + scale_shape_manual(values = c(24,25)) +
  scale_fill_manual(values=wes_palette(n=5, name="Cavalcanti1")) +
  scale_color_manual(values=wes_palette(n=5, name="Cavalcanti1")) +
    geom_vline(xintercept = 0, linetype=2) +
    geom_hline(yintercept = 0, linetype=2) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank()) +
    labs(fill = "Farm", shape= "Drought", color="Farm") + 
  xlab("PC1 (36% variance)") + ylab("PC2 (19% variance)") +
  theme(axis.text = element_text(size = 12)) + 
    theme(axis.title = element_text(size = 12)) +
    theme(legend.title = element_text(size = 12)) + 
    theme(legend.text = element_text(size = 12))

```

## Spain

```{r ES, message=FALSE, warning=FALSE}

es_meta <- sampledata %>%
  filter(Country == "ES")
es_meta$Farm <- as.factor(es_meta$Farm)

es_ID <- EZ_ID_dgelist[,colnames(EZ_ID_dgelist) %in% rownames(es_meta)]

set.seed(6359)
es_dd <- DESeqDataSetFromMatrix(countData = es_ID,
                              colData = es_meta[match(colnames(es_ID), rownames(es_meta)), ],
                              design= ~  Treatment + Farm)

es_vst <- varianceStabilizingTransformation(es_dd)

p1_ES_farm <- plotPCA(es_vst, "Farm")
p2_ES_treat <- plotPCA(es_vst, "Treatment")
p3_ES_time <- plotPCA(es_vst, "Time")
plot_grid(p1_ES_farm, p2_ES_treat, p3_ES_time)

#PERMANOVA
es_vst_counts <- assay(es_vst)
es_vst_counts <- t(es_vst_counts)
set.seed(4364)
adonis(es_vst_counts~ Treatment + Time, 
       data=es_meta, strata=es_meta$Farm, permutations=999)

```
***Drought effect, minor time effect***
***Differences between farms***

### plot PCA

```{r plot PCA ES}

ntop <- 500
rv <- rowVars(assay(es_vst))
select <- order(rv, decreasing = TRUE)[seq_len(min(ntop, length(rv)))]
mat <- t( assay(es_vst)[select, ] )

pca_spain<-prcomp(mat)

eigs <- pca_spain$sdev^2
eigs[1] / sum(eigs)
eigs[2] / sum(eigs)

PCA_ES_ID <- data.frame(pca_spain$x, Farm=es_meta$Farm, Time=es_meta$Time, Treat=es_meta$Treatment, Cou_Treat=es_meta$Cou_Treat)

#drought-time
ggplot() + geom_point(data=PCA_ES_ID, mapping=aes(x=PC1, y=PC2, color = Treat, shape=Time), size = 10,alpha=0.8) +
geom_vline(xintercept = 0, linetype=2) +
geom_hline(yintercept = 0, linetype=2) +
theme_bw() +
theme(panel.grid.major = element_blank(),
panel.grid.minor = element_blank()) +
scale_color_manual(values=c("#8f8f8f","#000000")) + labs(colour = "Drought", shape= "Sampling time") + guides(color = guide_legend(order = 1),
shape = guide_legend(order = 2)) +
theme(axis.text = element_text(size = 12)) +
theme(axis.title = element_text(size = 12)) +
theme(legend.title = element_text(size = 12)) +
theme(legend.text = element_text(size = 12))

#farm-drought
ggplot(PCA_ES_ID, aes(x=PC1, y=PC2, color = Farm, shape=Treat)) + geom_text_repel(aes(label=Time),max.overlaps = Inf, point.size = 8, segment.color=NA, show.legend = F,size=6) +
  geom_point(aes(fill = Farm, shape = Treat),size = 10,alpha=0.8) + scale_shape_manual(values = c(24,25)) +
  scale_fill_manual(values=wes_palette(n=5, name="Cavalcanti1")) +
  scale_color_manual(values=wes_palette(n=5, name="Cavalcanti1")) +
    geom_vline(xintercept = 0, linetype=2) +
    geom_hline(yintercept = 0, linetype=2) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank()) +
    labs(fill = "Farm", shape= "Drought", color="Farm") + 
  xlab("PC1 (26% variance)") + ylab("PC2 (19% variance)") +
  theme(axis.text = element_text(size = 12)) + 
    theme(axis.title = element_text(size = 12)) +
    theme(legend.title = element_text(size = 12)) + 
    theme(legend.text = element_text(size = 12))
  
#ggsave("../figures/PCA_ES_enzID.jpeg", width = 20, height = 15, units = "cm")

```

### batch ES
+ correction on farms

```{r ES batch correction, message=F, warning=F}

es_vst_remFP <- es_vst
assay(es_vst_remFP) <- limma::removeBatchEffect(assay(es_vst_remFP), es_vst_remFP$Farm)

p1_ES_FP_corr <- plotPCA(es_vst_remFP, "Farm")
p2_ES_treat_corr <- plotPCA(es_vst_remFP, "Treatment")
plot_grid(p1_ES_FP_corr, p2_ES_treat_corr)

#PERMANOVA 
es_vst_remFP_df <- as.data.frame(assay(es_vst_remFP))
es_vst_remFP_df <- t(es_vst_remFP_df)

set.seed(4364)
adonis(es_vst_remFP_df ~ Treatment, 
                     data=es_meta, permutations=999)

```
***Drought effect***

### plot batch ES

```{r ES batch plot PCA ES}

ntop <- 500
rv <- rowVars(assay(es_vst_remFP))
select <- order(rv, decreasing = TRUE)[seq_len(min(ntop, length(rv)))]
mat <- t( assay(es_vst_remFP)[select, ] )

pca_spain_batch<-prcomp(mat)

eigs <- pca_spain_batch$sdev^2
eigs[1] / sum(eigs)
eigs[2] / sum(eigs)

PCA_ES_ID_batch <- data.frame(pca_spain_batch$x, Farm=es_meta$Farm, Time=es_meta$Time, Treat=es_meta$Treatment, Cou_Treat=es_meta$Cou_Treat)

#drought-time
ggplot() + geom_point(data=PCA_ES_ID_batch, mapping=aes(x=PC1, y=PC2, color = Treat, shape=Time), size = 10,alpha=0.8) +
geom_vline(xintercept = 0, linetype=2) +
geom_hline(yintercept = 0, linetype=2) +
theme_bw() +
theme(panel.grid.major = element_blank(),
panel.grid.minor = element_blank()) +
scale_color_manual(values=c("#8f8f8f","#000000")) + labs(colour = "Drought", shape= "Sampling time") + guides(color = guide_legend(order = 1),
shape = guide_legend(order = 2)) +
theme(axis.text = element_text(size = 12)) +
theme(axis.title = element_text(size = 12)) +
theme(legend.title = element_text(size = 12)) +
theme(legend.text = element_text(size = 12))

#farm-drought
ggplot(PCA_ES_ID_batch, aes(x=PC1, y=PC2, color = Farm, shape=Treat)) + geom_text_repel(aes(label=Time),max.overlaps = Inf, point.size = 8, segment.color=NA, show.legend = F,size=6) +
  geom_point(aes(fill = Farm, shape = Treat),size = 10,alpha=0.8) + scale_shape_manual(values = c(24,25)) +
  scale_fill_manual(values=wes_palette(n=5, name="Cavalcanti1")) +
  scale_color_manual(values=wes_palette(n=5, name="Cavalcanti1")) +
    geom_vline(xintercept = 0, linetype=2) +
    geom_hline(yintercept = 0, linetype=2) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank()) +
    labs(fill = "Farm", shape= "Drought", color="Farm") + 
  xlab("PC1 (35% variance)") + ylab("PC2 (19% variance)") +
  theme(axis.text = element_text(size = 12)) + 
    theme(axis.title = element_text(size = 12)) +
    theme(legend.title = element_text(size = 12)) + 
    theme(legend.text = element_text(size = 12))
  
ggsave("../figures/PCA_ES_enzID_time_batch.jpeg", width = 20, height = 15, units = "cm")

```



```{r info}
sessionInfo()
```